---
title: "Permutations and Combinations"
output: html_notebook
---

## For a set of characters, work out the permutations 

My first attempt used a package called `gtools`. However, it produced incorrect results so I will try something else.
The `R` syntax for this is:
`permutations(n, r, v=1:n, set=TRUE, repeats.allowed=FALSE)`
`combinations(n, r, v=1:n, set=TRUE, repeats.allowed=FALSE)`

```{r}
library(stringr)
library(gtools)

# create a vector
players <- c("Baggio", "Pele", "Zidane", "Maradona", "Puskas", "Muller", "Yashin", "Beckenbauer", "Charlton", "Barnes", "Messi", "Ronaldo")
num <- length(players)

#fives <- permutations(num, 5, repeats.allowed=FALSE)

#fives

#str_glue("There are {fives} permutations.")


#permutations(n, r, v=1:n, set=TRUE, repeats.allowed=FALSE)
#str_glue("There are {perms_reps} permutations, allowing for repetitions.")

#THIS ATTEMPT DID NOT WORK. I SEARCHED ONLINE, BUT HAVE NOT FOUND A LIKELY PACKAGE THAT CAN PRODUCE THE CALCULATIONS I WANT. SO, I'LL MAKE MY OWN.

```


## Permutations, using my own function (no repetitions)

The mathematical formula is:

P=n!

To get the number of permutations for a subset. For instance, how many two-letter combinations can you make from the alphabet?

nPr = n! / (n-r)!

_where n is the number of items in the population_

_where r is the size of the subset_

_where P is the number of permutations_

```{r}
# create function
permutator <- function(populationSize, targetSize=populationSize) {
  if (populationSize==targetSize) {
    perms <- factorial(populationSize)
  } else {
    perms <- (factorial(populationSize)) / (factorial(populationSize-targetSize))
  }
  return(perms)
}

fives <- permutator(num)
fives 

abcd <- c("A", "B", "C", "D")
abcdLen <- length((abcd))
quad <- permutator(abcdLen)
quad

duo <- permutator(abcdLen, 2)
duo

```
## Permutations with repetitions
This one is easy. It's just n to the power of r.
```{r}
reps <- abcdLen ^ 3
reps
```



## For a set of characters, work out the combinations 

nCr = n! / r! * (n-r)!

_where C is combination_ 



```{r}
#create a function
combinator <- function(populationSize, targetSize) {
  combs <- (factorial(populationSize)) / (factorial(targetSize) * factorial(populationSize-targetSize))
  return(combs)
}

#use the function
playerCombs <- combinator(num, 5)
formatted <- format(playerCombs, big.mark=",")
formatted

```

## Conbinations with repetitions

n+r-1 C r = (n+r-1)! / r! * (n-1)!
```{r}
#create a function
rep_combinator <- function(populationSize, targetSize) {
  combs <- (factorial(populationSize+targetSize+1)) / (factorial(targetSize) * factorial(populationSize-1))
  return(combs)
}

#use the function
repCombs <- rep_combinator(num, 5)
rep_formatted <- format(repCombs, big.mark=",", scientific = FALSE)
rep_formatted
```


